Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-07-14 Prelim/google_research_overview.tex
uploaded editor comments

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\begin{document}
	\title{Efficient and Scalable Bayesian Bipartite Matching through Fast Beta Linkage (\texttt{fabl})}
	%\subtitle{Using Beamer}
	\author{Brian Kundinger, Jerome Reiter, Rebecca Steorts}
	\institute{Duke University}
	\date{\today}

@@

\begin{frame}{Linkage through Comparison Vectors}
	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
\begin{center}
	$Z_j = \begin{cases} 
		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
	\end{cases}$
\end{center}
\end{frame}

\begin{frame}{Linkage through Comparison Vectors}
	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
\end{frame}

@@

	Model specification allows for \textbf{hashing} of comparison vectors, \textbf{distributed} computing, and \textbf{storage efficient indexing (SEI)}
\end{frame}

\begin{frame}{Hashing}
		\begin{columns}
		\begin{column}{0.6\textwidth}
			\begin{itemize}
				\item<1-> Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records (Enamorado et al 2019).
				\begin{itemize}
					\item<2-> $L = \{3, 3, 2, 2\}$ implies 36 unique patterns 
				\end{itemize} 
				\item<3-> When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p.$
				\item<4-> Allows us to compute sufficient statistics and reduce computational complexity from $O(n_A \times n_B)$ to $O(P \times n_B)$
			\end{itemize}
		\end{column}
		\begin{column}{0.38\textwidth}
			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics/unique_patterns.png}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Hashing}
	\centering
	\includegraphics<1>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide17.png}
	\includegraphics<2>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide18.png}
	\includegraphics<3>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide19.png}
	\includegraphics<4>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide20.png}
	\includegraphics<5>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide21.png}	
\end{frame}

\begin{frame}{Hashing to Scale to Large Linkage Tasks}
	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide23.png}
	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide24.png}
	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide25.png}
	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide26.png}
	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide27.png}
\end{frame}

\begin{frame}{Hashing for Speed (Efficient Gibbs Sampling)}
	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide37.png}
	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide38.png}
	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide39.png}
	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide40.png}
	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide41.png}
	\includegraphics<6>[width = \textwidth, height = .6\textwidth ]{graphics/Slide42.png}
\end{frame}

@@

\begin{frame}{Extensions and Future Directions}
	\begin{itemize}
		\item Linkage when reliability of information and rates of matching differs by subgroup in the data

		\item Linkage when there are duplicates within datasets

		\item Linkage over blocked data (allows for much larger linkage tasks)
	\end{itemize}
\end{frame}

\section{Appendix}

\begin{frame}{Storage Efficient Indexing}
	\begin{itemize}
		\item Storing $n_A \times n_B$ hashed values burdensome for large tasks
		\item Overwhelming majority of record pairs show nonagreement
		\item<3> Correct counts for calculations stored in $\mathcal{H}$
	\end{itemize}
	\includegraphics<1>[width = \textwidth]{graphics/before_SEI.png}
	\includegraphics<2->[width = \textwidth]{graphics/after_SEI2.png}
\end{frame}

%		\item Reduce data to sufficient statistics
%		\begin{itemize}
%		\item $r_{p_j} = \{i \;|\, (i, j) \in h_p\}$
%		\item $H_{p_j} = ||r_{p_j}||$
%		\item $H_p = \sum_{j} H_{p_j}$
%		\end{itemize}
%		
%		\item Run Gibbs sampler at level of agreement patterns, not record pairs
%		\begin{itemize}
%		\item Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
%		\item Use number of matches for each pattern to update $m$ and $u$
%		\item Back fill record labels at the end through $r_{p_j}$
%	\end{itemize}
%
%		\item Reduces computational complexity from $O(n_A \times n_B \times F)$ to $O(P \times n_B \times F)$

>>>>

\documentclass{beamer}
\usetheme{Rochester}
\addtobeamertemplate{navigation symbols}{}{ \hspace{1em}    \usebeamerfont{footline}%
	\insertframenumber / \inserttotalframenumber }
\begin{document}
	\title{Efficient and Scalable Bayesian Bipartite Matching through Fast Beta Linkage (\texttt{fabl})}
	\subtitle{Google Version}
	\author{Brian Kundinger, Jerome Reiter, Rebecca Steorts}
	\institute{Duke University}
	\date{\today}

@@

%\begin{frame}{Linkage through Comparison Vectors}
%	Represent linkage structure through vector $\mathbf{Z} = \{Z_1, \ldots, Z_{n_B}\}$, where
%\begin{center}
%	$Z_j = \begin{cases} 
%		i,  & \text{if records } i\in A \text{ and } j\in B \text{ match}; \\
%		n_A + 1,  & \text{if record } j\in B \text{ has no match in } A; \\
%	\end{cases}$
%\end{center}
%\end{frame}
%
%\begin{frame}{Linkage through Comparison Vectors}
%	\includegraphics[width = \textwidth, height = .6\textwidth ]{graphics/Slide6.png}
%\end{frame}

@@

	Note: the first three lines are common to many record linkage models, and bear similarities to a larger family of \textit{latent class models.} The last two lines are the newly developed prior distribution for the set of matches.  
\end{frame}

%\begin{frame}{Hashing}
%		\begin{columns}
%		\begin{column}{0.6\textwidth}
%			\begin{itemize}
%				\item<1-> Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records (Enamorado et al 2019).
%				\begin{itemize}
%					\item<2-> $L = \{3, 3, 2, 2\}$ implies 36 unique patterns 
%				\end{itemize} 
%				\item<3-> When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p.$
%				\item<4-> Allows us to compute sufficient statistics and reduce computational complexity from $O(n_A \times n_B)$ to $O(P \times n_B)$
%			\end{itemize}
%		\end{column}
%		\begin{column}{0.38\textwidth}
%			\includegraphics<2->[width = \textwidth, height = 1.2\textwidth ]{graphics/unique_patterns.png}
%		\end{column}
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Hashing}
%	\centering
%	\includegraphics<1>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide17.png}
%	\includegraphics<2>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide18.png}
%	\includegraphics<3>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide19.png}
%	\includegraphics<4>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide20.png}
%	\includegraphics<5>[width = 1.1\textwidth, height = .6\textwidth ]{graphics/Slide21.png}	
%\end{frame}
%
%\begin{frame}{Hashing to Scale to Large Linkage Tasks}
%	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide23.png}
%	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide24.png}
%	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide25.png}
%	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide26.png}
%	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide27.png}
%\end{frame}
%
%\begin{frame}{Hashing for Speed (Efficient Gibbs Sampling)}
%	\includegraphics<1>[width = \textwidth, height = .6\textwidth ]{graphics/Slide37.png}
%	\includegraphics<2>[width = \textwidth, height = .6\textwidth ]{graphics/Slide38.png}
%	\includegraphics<3>[width = \textwidth, height = .6\textwidth ]{graphics/Slide39.png}
%	\includegraphics<4>[width = \textwidth, height = .6\textwidth ]{graphics/Slide40.png}
%	\includegraphics<5>[width = \textwidth, height = .6\textwidth ]{graphics/Slide41.png}
%	\includegraphics<6>[width = \textwidth, height = .6\textwidth ]{graphics/Slide42.png}
%\end{frame}

@@

\begin{frame}{Glaring Questions}
	\begin{itemize}
		\item Can you do record linkage with arbitrary amounts of duplicated files?
		\begin{itemize}
			\item<2-> Yes! See Aleshin-Guendel and Sadinle (2021)
		\end{itemize}

		\item Isn't $n_A \times n_B$ record pairs computationally infeasible?
		\begin{itemize}
			\item<3-> Yes! My work with the DNC is forthcoming, Kundinger, Wortman, Reiter, and Steorts (2022)
		\end{itemize}

		\item Can you account situations when the reliability of information differs throughout the data?
		\begin{itemize}
			\item<4-> Yes! I have a heirarchical model in the works. Hopefully Kundinger et al (2023)
		\end{itemize}

		\item Can you do record linkage on the records themselves, rather than transforming to comparison vectors?
		\begin{itemize}
			\item<5-> Yes! See \texttt{d-blink} package from Marchant et al (2021)
		\end{itemize}

	\end{itemize}
\end{frame}

\section{Possible Relevance to Google}
\begin{frame}{Latent Class Analysis}
	The record linkage model presented is a special case of a "latent class model". In my work, we attempt to classify record pairs into two distinct classes: matching, and non-matching pairs.
	\linebreak

	One can imagine however, that we can look at articles instead of records, look at blocks of text instead of record pairs, and look at attributes of those blocks instead of comparison vectors. This may provide a strategy for \textbf{entirely unsupervised} byline detection.
	\linebreak

	Additionally, if model parameters are accurate, we can train the model on a small set of articles, and then apply those parameters out onto a larger set of articles, for nearly instantaneous latent class analysis. 
\end{frame}

### 2022-07-14 methods_paper/cover-letter/ba-cover-letter.tex
uploaded editor comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[top=3in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
                      % LOOKING THE WAY YOU WANT IT
	bottom=1in, left=1in, right=1in]{geometry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eso-pic}
\usepackage{graphicx}

@@

%\setlength\headheight{25em}
%\setlength\headsep{0em}
%\date{June 9, 2010}
\date{\raggedright November 29, 2021}

\begin{document}

\AddToShipoutPicture*{\BackgroundPic}
\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
\thispagestyle{empty}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
                % CHANGE THIS NUMBER TO GET THE FIRST PAGE HOW YOU WANT IT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

I am a PhD student at Duke University in the Department of Statistical Science. I am writing to submit my paper with my coauthors
Jerome Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.

Very often information about social entities is scattered across multiple databases.  Combining that information into one database can result in enormous benefits for analysis, resulting in richer and more reliable conclusions. The application areas touched by record linkage include voter registration, official statistics, human rights violations, patent inventions, health applications, among many others. Disciplines that utilize both applications and methods include statistics, biostatistics, computer science, machine learning, political science, and others. The goal of our review article is to review nearly almost all of entity resolution in such a way that it can be read and understood by as many communities as possible given that the current state of the literature is mathematical and overly technical, making the entry point difficult for those in the computational social sciences.

@@

To help the editorial office, we have listed a number of potential reviewers that do not have any conflicts with the authors:

Mauricio Sadinle, msadinle@uw.edu, University of Washington; Serge Aleshin-Guendel, aleshing@uw.edu, University of Washington; Yves Thibaudeau, Yves.Thibaudeau@census.gov, United States Census Bureau; Megan Price, meganp@hrdag.org, Human Rights Data Analysis Group; Ted Enamorado, ted@wustl.edu, University of Washington St. Louis; Nick Monath, nmonath@cs.umass.edu; Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au, Ben Rubinstein; benjamin.rubinstein@unimelb.edu.au; Neil Marchant; nmarchant@unimelb.edu.au; Andy Toole; Andrew.Toole@uspto.gov;  jph34@stat.duke.edu; Giacomo Zanella, zanella.gcm@gmail.com, Brunero Liseo, brunero.liseo@uniroma1.it, Andrea Tancredi, andrea.tancredi@uniroma1.it;
Peter Christen, peter.christen@anu.edu.au, Partha Lahiri, plahiri@umd.edu, Rebecca Nugent, rnugent@stat.cmu.edu, Kayla Frisoli, kaylafrisoli@gmail.com.

@@

 \closing{Sincerely,}

 Rebecca C. Steorts

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[top=1in,  % FIRST, CHANGE THIS NUMBER TO GET THE SECOND PAGE 
                      % LOOKING THE WAY YOU WANT IT
	bottom=1in, left=1in, right=1in]{geometry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{eso-pic}
%\usepackage{graphicx}
\usepackage{parskip}

@@

%\setlength\headheight{25em}
%\setlength\headsep{0em}
%\date{June 9, 2010}
%\date{\raggedright November 29, 2021}

\begin{document}
\date{\raggedright December 21, 2021}
%\AddToShipoutPicture*{\BackgroundPic}
%\signature{\vspace*{-4em}\includegraphics[scale=1]{sign.pdf}} 
%\thispagestyle{empty}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\vspace*{-.2in}  % AFTER YOU HAVE THE SECOND PAGE THE WAY YOU WANT IT,
                % CHANGE THIS NUMBER TO GET THE FIRST PAGE HOW YOU WANT IT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

I am a PhD student at Duke University in the Department of Statistical Science. I am writing to submit my paper with my coauthors
Jerome P. Reiter and Rebecca C. Steorts entitled ``Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage (fabl)" to \emph{Bayesian Analysis}.

Very often information about social entities is scattered across multiple databases.  Combining that information into one database can result in enormous benefits for analysis, resulting in richer and more reliable conclusions. The application areas touched by record linkage include voter registration, official statistics, human rights violations, patent inventions, health applications, among many others. Disciplines that utilize both applications and methods include statistics, biostatistics, computer science, machine learning, political science, and others.

@@

To help the editorial office, we have listed a number of potential reviewers:

\setlength{\parindent}{0mm}
\hangindent=10mm
\hangafter=0
Mauricio Sadinle, msadinle@uw.edu \newline
Serge Aleshin-Guendel, aleshing@uw.edu \newline
Yves Thibaudeau, Yves.Thibaudeau@census.gov \newline
Megan Price, meganp@hrdag.org \newline
Ted Enamorado, ted@wustl.edu \newline
Nick Monath, nmonath@cs.umass.edu \newline
Dinusha Vatsalan, Dinusha.Vatsalan@data61.csiro.au \newline
Ben Rubinstein, benjamin.rubinstein@unimelb.edu.au \newline
Neil Marchant, nmarchant@unimelb.edu.au \newline
Andy Toole, Andrew.Toole@uspto.gov \newline
Giacomo Zanella, zanella.gcm@gmail.com \newline
Brunero Liseo, brunero.liseo@uniroma1.it \newline
Andrea Tancredi, andrea.tancredi@uniroma1.it \newline
Peter Christen, peter.christen@anu.edu.au \newline
Partha Lahiri, plahiri@umd.edu \newline
Rebecca Nugent, rnugent@stat.cmu.edu \newline
Kayla Frisoli, kaylafrisoli@gmail.com \newline

@@

 \closing{Sincerely,}
 Brian Kundinger
